The Units of Permissionless Consensus: \\ Towards Mobile and Edge Computing

### 2022-10-11 theoretical-consensus/article/article.tex
add page 03

%% bare_jrnl.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

@@

%\usepackage[pdftex]{graphicx}

@@

In the classical Nakamoto consensus protocol, the generation of a block, to be proposed
for further network agreement, complies with the unit of computational work needed to
create, or rather find, a verifiable proof of the effort spent on creating the block \cite{nakamoto2008bitcoin}. 
This essentially requires brute forcing the finding of a cryptographic hash value for the
aggregation of the block information with a nonce such that this hash value satisfies
a difficulty threshold, which gets adjusted dynamically over time, to maintain the network overall 
requirement for the block generation interval \cite{survey-dist-consensus}.

Mitigation of Sybil attacks

\begin{algorithm}
\caption[short]{BlockGeneration}
\begin{algorithmic}[1]
  \Function {}{}
    \State $Block Header \gets Transaction \ Merkle \ Tree \ Root$
    \Indent
      \State $| \ Hash \ of \ the \ last \ Block$
      \State $| \ Timestamp$
      \State $| \ Other$
    \EndIndent

@@

>>>>

%% article.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

@@

\usepackage[pdftex]{graphicx}
\graphicspath{ {./images/} }

@@

In the classical Nakamoto consensus protocol, the generation of a block, to be proposed
for further network agreement, complies with the unit of computational work needed to
create, or rather find, a verifiable proof of the effort spent on creating the block \cite{nakamoto2008bitcoin}. 
This essentially requires brute forcing the search for a cryptographic hash value for the
aggregation of the block information with a nonce, such that this hash value satisfies
a difficulty threshold (see Algorithm \ref{alg:BlockGeneration}), which gets adjusted dynamically over time, to maintain the network overall 
requirement for the block generation interval \cite{survey-dist-consensus}.

There are a couple of particularities that need to be mentioned. First, the block generation
is of probablistic nature, as the nonce is a random value, and the hash function is a one-way function.
Allied to this, more hashing power is not easily obtained, which allows for tackling Sybil attacks,
despite the permissionless and pseudonymity nature of the network. Plus, the incentive mechanism
also plays a role in this regard, by naturally incentivizing honest participation.
The block generation interval, forged by the adjustment of the difficulty
threshold, is also of major importance because it reduces the fork probability and
allows for the timely propagation of proposed blocks, guaranteeing consensus
in the presence of an honest majority of the hashing power \cite{garay2015bitcoin}.
This probabilitic finality shifts the 1/3 BFT threshold to a 1/2 threshold.

\begin{algorithm}
\caption[short]{BlockGeneration}\label{alg:BlockGeneration}
\begin{algorithmic}[1]
  \Function {}{}
    \State $Block Header \gets Transaction \ Merkle \ Tree \ Root$
    \Indent
      \State $| \ Hash \ of \ the \ last \ Block$
      \State $| \ Timestamp$
      \State $| \ Other$
    \EndIndent

@@

\subsection{Trade-offs and trillema}

Despite the harmony of all the aforementioned clever characteristics,
the proof-of-work mechanism has some well known drawbacks. 
First, the block generation is computationally expensive, and so the
energy consumption associated with it is also high. This is caused by the
dynamically adjusted difficulty threshold, which is a function of the block 
generation interval. To maintain the 10 minutes interval, the hashing difficulty
increases, and so too the hashing power and the energy needed for it. 
However, reducing the block generation interval would increase the
probability of forks, which results in a more frequent waste of useful work,
loosening the security of the network. Increasing the block size would also
have the same security effect due to the increased propagation delay.

As pointed in the literature, the design of these consensus mechanisms
shall aim for a protocular choice between a set of properties that form a trillema (see Fig. \ref{fig:trillema}): 
Security, Scalability, and Decentralization \cite{survey-dist-consensus}.
Summarizing, ...

\begin{figure}[h]
  \centering
  \includegraphics[width=\columnwidth]{trillema}
  \caption{From \cite{survey-dist-consensus}, the blockchain consensus trillema.}
  \label{fig:trillema}
\end{figure}
